---
title: "Data Visualization HW2"
author: "Arman Manukyan"
date: "2025-02-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
if (!require("car")) install.packages("car")
library(car)
if (!require("lsmeans")) install.packages("lsmeans")
library(lsmeans)
if (!require("dplyr")) install.packages("dplyr")
library(dplyr)
if (!require("tidyverse")) install.packages("tidyverse")
library(tidyverse)
if (!require("ggplot2")) install.packages("ggplot2")
library(ggplot2)
if (!require("emmeans")) install.packages("emmeans")
library(emmeans)
library(ggthemes)
library(knitr)
library(reticulate)
library(GGally)
library(dplyr)
library(ggplot2)
library(ggthemes)
library(knitr)
library(reticulate)
library(GGally)
library(reshape2)
library(corrplot)
```

Part 3: Use the datasets provided to create graphs
Overview of the datasets.
The “lung_cancer_prediction_dataset.csv” provides valuable insights into lung cancer cases, risk
factors, smoking trends, and healthcare access across 25 of the world's most populated
countries. It includes 220,632 individuals with details on their age, gender, smoking history,
cancer diagnosis, environmental exposure, and survival rates. The dataset is useful for medical
research, predictive modeling, and policymaking to understand lung cancer patterns globally.
source


The “global_air_pollution_dataset.csv” provides information about the contamination of the
indoor or outdoor environment by any chemical, physical or biological agent that modifies the
natural characteristics of the atmosphere. It includes 23,463 rows with data presented in the
following columns: Country, City, AQI Value, AQI Category, CO AQI Category, Ozone AQI Value,
Ozone AQI Category, NO2 AQI Value, NO2 AQI Category, PM2.5 AQI Value, PM2.5 AQI Category.
source

```{r Data lung_cancer_prediction_dataset.csv  }
data <- read.csv("lung_cancer_prediction_dataset.csv")
head(data)

```


```{r 1. Create a Boxplot of Lung Cancer Deaths Distribution.(R)}
ggplot(data, aes(y=Annual_Lung_Cancer_Deaths)) + 
  geom_boxplot(fill = 'darkorange1') +
  labs(title = "Boxplot of Lung Cancer Deaths Distribution ", y = "Annual Lung Cancer Deaths",color = "red", size = 10, face = "bold" ,subtitle="Lung Cancer Deaths")+
theme_bw()+
   theme(plot.title = element_text(color = "Black", size = 12, face = "bold"),
         axis.title.y = element_text(color = "darkorange", size = 12, face = "bold"),
         plot.subtitle = element_text(color = "darkorange4"))
```

```{r 1. Create a Boxplot of Lung Cancer Deaths Distribution.(R)}
ggplot(data, aes(y=Annual_Lung_Cancer_Deaths)) + 
  geom_boxplot(fill = 'darkorange1') +
  labs(title = "Boxplot of Lung Cancer Deaths Distribution ", y = "Annual Lung Cancer Deaths",color = "red", size = 10, face = "bold" ,subtitle="Lung Cancer Deaths")+
theme_bw()+
   theme(plot.title = element_text(color = "darkorange3", size = 12, face = "bold"),
         axis.title.x = element_text(color = "darkorange", size = 12, face = "bold"),
         plot.subtitle = element_text(color = "darkorange4"))+
    coord_flip()

```

```{r Data global_air_pollution_dataset.csv  }
data1 <- read.csv("global_air_pollution_dataset.csv")
head(data1)
summary(data1)

```
```{r,2. Create a Histogram of PM2.5 AQI Values.)}
ggplot(data1, aes(x=PM2.5_AQI_Value)) + 
  geom_histogram(bins=25,color = 'gold3', fill="darkseagreen4") +
  labs(title = "Histogram of PM2.5 AQI Values", x = "PM2.5_AQI_Value",subtitle = "PM2.5 AQI Values", y =NULL)+
  theme_minimal()+
  theme(plot.title = element_text(color = "firebrick4", size = 12, face = "italic"),
         axis.title.x = element_text(color = "darkslategray4", size = 12, face = "bold"),
         plot.subtitle = element_text(color = "darkorange4"))
```

Create a Density Plot of the Lung Cancer Mortality Rate. (Python & R)

```{r Create a Density Plot of the Lung Cancer Mortality Rate. (Python & R)}
ggplot(data, aes(x=Mortality_Rate)) + 
  geom_density(alpha = 0.5,color="red", fill="blue", lwd=1) +
  labs(title = "Density Plot of the Lung Cancer Mortality Rate", x = "Mortality_Rate",y="Density")+
theme_bw()+
  theme(panel.background = element_rect(fill = "steelblue", colour = NA, size = 0.5, linetype = "solid"))
```

Create a Scatter Plot by generating 100 random values from both the normal and logistic
distributions. The points should be brown and use theme_solarized with argument light
set to false.

```{r  Scatter Plot}
normal <- rnorm(100, mean = 0, sd = 1)
logistic <- rlogis(100, location = 0, scale = 1)

df <- data.frame(normal, logistic)

ggplot(df, aes(x = normal, y = logistic)) +
  geom_point(color = "brown") +
  theme_solarized(light = FALSE) +
  labs(title ="Scatter Plot of Normal vs Logistic Distribution", subtitle ="Normal vs Logistic Distribution", x="Normal Distribution Values",y="Logistic Distribution Values")+
  theme_solarized(light = FALSE)+
 theme(plot.title = element_text(color = "brown", size = 14, hjust = 0.5),
       plot.subtitle = element_text(hjust = 0.5))


```

Part 4: Recreate the following graphs


```{r}
pm25_agg <- aggregate(PM2.5_AQI_Value ~ Country, data1, FUN = mean)
lung_agg <- aggregate(Annual_Lung_Cancer_Deaths ~ Country, data, FUN = sum)
merged_data <- merge(pm25_agg, lung_agg, by = "Country")

ggplot(merged_data, aes(x = PM2.5_AQI_Value, y = Annual_Lung_Cancer_Deaths, color = Country )) +
  geom_point() +
  geom_text(aes(label = ifelse(Annual_Lung_Cancer_Deaths > 15000 | PM2.5_AQI_Value >100, Country, "")), 
            vjust = -0.5, hjust = 0.5, size = 2, fontface = "bold") +
  labs(title = "PM2.5 AQI vs. Annual Lung Cancer Deaths", x = "PM2.5 AQI Value", y = "Annual Lung Cancer Deaths", color = 'Annual Lung Cancer Deaths') +
  theme_bw()+
 theme(plot.title = element_text(hjust = 0.5, face = "bold", colour = "brown"),
      legend.position="right",
      legend.title = element_text(size = 12, face = "bold"),
      axis.text.x = element_text(color = 'blue'))

```


3. Use the ggplot2 package for this graph. (Hint: use geom_jitter since y axis contains
categorical data, also use the following colors: #5469f1 , #d554f1)

```{r}

data_filtered <- data[data$Cancer_Stage != 'None', ]
ggplot(data_filtered, aes(x = Years_of_Smoking, y = factor(Cancer_Stage), color = Gender,shape=Gender)) +
  geom_jitter(width = 0.2, height = 0.2, alpha = 0.6) +
  facet_wrap(~ Gender) +
  scale_color_manual(values = c("#d554f1", "#5469f1")) +
  scale_shape_manual(values = c("Female" = 17, "Male" = 16)) +
  labs(title = "Lung Cancer Stage vs. Smoking Years",
       subtitle = "Comparison by Gender",x = "Years of Smoking",y = "Cancer Stage") +
  theme_minimal() +
  theme(legend.position = "bottom", legend.direction = "horizontal",
        plot.title = element_text(face="bold"),
        strip.text = element_text(face = "bold", size = 12))

```


4.Use the ggplot2 package for this graph. (Hint: use scale_fill_viridis_d(option = "plasma"
to get the same colors)
```{r}
country  <- c("Brazil", "Germany", "India", "Italy", "Russian Federation", "United States of America")
filtered_data <- data1[data1$Country %in% country, ]
ggplot(filtered_data, aes(x = PM2.5_AQI_Value, fill = Country)) +
  geom_histogram(bin=30, binwidth = 10, color = "black") +
  labs(title = "PM2.5 AQI Distribution Across Countries", x = "PM2.5_AQI_Value", y = "Frequency",subtitle = "Comparison of Air Pollution levels", color = "Country") + theme_bw() + facet_wrap(~ Country, scales = "free")+ scale_fill_viridis_d(option = "plasma")+
  theme(legend.position = "bottom", legend.direction = "horizontal", strip.background = element_blank(),
        strip.text = element_text(face = "bold", size = 12),
        plot.background = element_rect(colour = "black", fill=, size=1))
  
```






















