---
title: "Homework N1 Crime"
author: "Arman Manukyan LATEST"
date: "2025-02-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = FALSE)
options(tinytex.verbose = TRUE)

options(scipen = 999)

library(dplyr)
library(ggplot2)
library(ggthemes)
library(knitr)
library(reticulate)
library(lubridate)
library(GGally)
```


```{r Loading a data file}
# Load the data file interactively, keeping the original code style
crime.data <- read.table(file.choose(), header = FALSE, sep = ",", quote = "\"", 
                      stringsAsFactors = FALSE, fileEncoding = "ISO-8859-1")
# Assign column name since the dataset lacks one
colnames(crime.data) <- c("Value")

# Display first few rows and summary
head(crime.data)
summary(crime.data)

# Open the data frame viewer
View(crime.data)

```
Identify columns with missing values and their respective counts. Drop columns where more than
50% of the data is missing (store this version as a new dataset).

```{r Identify and Remove Missing Values}
# Identify columns with missing values and their respective counts
missing_counts <- colSums(is.na(crime.data))

# Display missing values count per column
print(missing_counts)

# Define the threshold (50% missing values)
threshold <- 0.5 * nrow(crime.data)

# Drop columns with more than 50% missing values
crime.cleaned <- crime.data[, colSums(is.na(crime.data)) <= threshold]

# Save the cleaned dataset as a CSV file
write.csv(crime.cleaned, "crime_cleaned.csv", row.names = FALSE)

# Provide a download link
cat("Download the cleaned dataset: [crime_cleaned.csv](crime_cleaned.csv)")

```
Convert the DATE OCC column to a datetime format. Extract the year, month, and day into separate
columns. Create a new column for the hour using the TIME OCC column


```{r Convert and Extract Date-Time Components}
# Ensure the first row is treated as column names
colnames(crime.cleaned) <- as.character(crime.cleaned[1, ])
crime.cleaned <- crime.cleaned[-1, ]

# Convert 'DATE OCC' column to a date format using as.Date()
crime.cleaned$`DATE OCC` <- as.Date(crime.cleaned$`DATE OCC`, format = "%m/%d/%Y")

# Extract year, month, and day into separate columns
crime.cleaned$Year <- format(crime.cleaned$`DATE OCC`, "%Y")
crime.cleaned$Month <- format(crime.cleaned$`DATE OCC`, "%m")
crime.cleaned$Day <- format(crime.cleaned$`DATE OCC`, "%d")

# Convert 'TIME OCC' column to an hour column
crime.cleaned$Hour <- as.numeric(substr(crime.cleaned$`TIME OCC`, 1, 2))

# Display the first few rows of the updated dataset
head(crime.cleaned)

# View the updated dataset
View(crime.cleaned)

```
Filter the dataset for crimes that occurred in 2023. Further filter crimes with the description
BURGLARY in the Crm Cd Desc column.
```{r Filter Data for 2023 and BURGLARY Crimes}
# Ensure 'Year' column is numeric for filtering
crime.cleaned$Year <- as.numeric(crime.cleaned$Year)

# Filter dataset for crimes that occurred in 2023
crime.2023 <- subset(crime.cleaned, Year == 2023)

# Further filter for crimes with description 'BURGLARY' in the 'Crm Cd Desc' column
crime.burglary.2023 <- subset(crime.2023, `Crm Cd Desc` == "BURGLARY")

# Display the first few rows of the filtered dataset
head(crime.burglary.2023)

# View the filtered dataset
View(crime.burglary.2023)

```
Group the data by AREA NAME and calculate the total number of crimes and the average victim age.
Sort the results by total crimes in descending order.

```{r Group Data by AREA NAME and Analyze}

# Ensure 'Vict Age' is numeric for calculations
crime.cleaned$`Vict Age` <- as.numeric(crime.cleaned$`Vict Age`)

# Group by AREA NAME, calculate total number of crimes and average victim age
crime_summary <- crime.cleaned %>%
  group_by(`AREA NAME`) %>%
  summarise(Total_Crimes = n(),
            Avg_Victim_Age = mean(`Vict Age`, na.rm = TRUE)) %>%
  arrange(desc(Total_Crimes))  # Sort by total crimes in descending order

# Display the summary dataset
head(crime_summary)

# View the grouped summary dataset
View(crime_summary)

```

Part 3: Further Exploration (R only) 5 points per problem
Group the data by Month and count the number of crimes.
2. Count the number of crimes where a weapon was used (Weapon Used Cd is not null).
3. Group the data by Premis Desc and count the number of crimes.

```{r Group and Count Crimes by Month, Weapon Used, and Premises}
# 1. Group the data by Month and count the number of crimes
crime_by_month <- crime.cleaned %>%
  group_by(Month) %>%
  summarise(Total_Crimes = n()) %>%
  arrange(Month)  # Ensure sorting by month order

# Display the first few rows
head(crime_by_month)

# View the grouped dataset
View(crime_by_month)


# 2. Count the number of crimes where a weapon was used (Weapon Used Cd is not null)
weapon_used_crimes <- crime.cleaned %>%
  filter(!is.na(`Weapon Used Cd`)) %>%
  summarise(Total_Weapon_Crimes = n())

# Display the result
print(weapon_used_crimes)
View(weapon_used_crimes)

# 3. Group the data by Premis Desc and count the number of crimes
crime_by_premises <- crime.cleaned %>%
  group_by(`Premis Desc`) %>%
  summarise(Total_Crimes = n()) %>%
  arrange(desc(Total_Crimes))  # Sort in descending order

# Display the first few rows
head(crime_by_premises)

# View the grouped dataset
View(crime_by_premises)

```

Part 4: Advanced Analysis (Python & R) 20 points per problem: 10 for Python and 10 for R
1. Create a new column, Severity Score, based on the following rules:
• Assign 5 points if a weapon was used (Weapon Used Cd is not null).
• Assign 3 points for crimes under BURGLARY.
• Assign 1 point for all other crimes.
• Group by AREA NAME and find the total severity score for each area.

```{r Calculate Severity Score and Group by AREA NAME}
# Ensure Weapon Used Cd is treated correctly for missing values
crime.cleaned$`Weapon Used Cd` <- as.character(crime.cleaned$`Weapon Used Cd`)

# Assign default Severity Score of 1
crime.cleaned$Severity_Score <- 1

# Assign 3 points for BURGLARY crimes
crime.cleaned$Severity_Score[crime.cleaned$`Crm Cd Desc` == "BURGLARY"] <- 
  crime.cleaned$Severity_Score[crime.cleaned$`Crm Cd Desc` == "BURGLARY"] + 3

# Assign 5 points if a weapon was used (Weapon Used Cd is not null)
crime.cleaned$Severity_Score[!is.na(crime.cleaned$`Weapon Used Cd`)] <- 
  crime.cleaned$Severity_Score[!is.na(crime.cleaned$`Weapon Used Cd`)] + 5

# Group by AREA NAME and calculate total severity score
severity_by_area <- crime.cleaned %>%
  group_by(`AREA NAME`) %>%
  summarise(Total_Severity_Score = sum(Severity_Score)) %>%
  arrange(desc(Total_Severity_Score))
# Display results
head(severity_by_area)
# View dataset
View(severity_by_area)

```

Bonus Part: 5 points
Use the LAT and LON columns to identify crimes that occurred within a specific latitude-longitude bounding
box (e.g., downtown area).

```{r Filter Crimes by Latitude and Longitude}
# Convert LAT and LON to numeric (if necessary)
crime.cleaned$LAT <- as.numeric(crime.cleaned$LAT)
crime.cleaned$LON <- as.numeric(crime.cleaned$LON)

# Define the bounding box (example: downtown area)
lat_min <- 34.04
lat_max <- 34.07
lon_min <- -118.27
lon_max <- -118.23

# Filter the dataset based on the bounding box
crime_downtown <- crime.cleaned %>%
  filter(LAT >= lat_min & LAT <= lat_max & LON >= lon_min & LON <= lon_max)

# Display results
head(crime_downtown)

# View the dataset
View(crime_downtown)

```


















