---
title: "Data Visualization"
author: "Arman Manukyan, Mariam Manukyan"
date: "2025-04-19"
Professor: " Gurgen Hovakimyan"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    editor_options: 
      chunk_output_type: console
    theme:
      version: 4
      bg: "#101010"
      fg: "#FDF7F7" 
      primary: "#ED79F9"
      navbar-bg: "#3ADAC6"
      base_font: 
        google: Prompt
      heading_font:
        google: Sen
      code_font:
        google: 
          # arguments to sass::font_google() 
          family: JetBrains Mono
          local: false
---

```{r setup, include=FALSE}
library(flexdashboard)
library(ggplot2)
library(dplyr)
library(purrr)
library(plotly)
library(DT)
library(moments)
library(magrittr)
library(ggthemes)
library(viridis)
library(RColorBrewer)
library(data.table)
library(summarytools)
library(parameters)
library(jsonlite)
library(tidyr)

```

Column {data-width=650}



# Load the data file 

```{r Loading a data file}
# Let the user browse for the file
file_path <- "fullData.csv"

# Read the data into the variable 'fulldata'
fulldata <- read.csv(file_path)

# Display a message to confirm successful reading
cat("Data successfully read into the variable 'fulldata'.\n")

```

-----------------------------------------------------------------------

### CHART 1

```{r}

# Load necessary libraries
library(dplyr)
library(ggplot2)

# Process the data: Count reading frequency by age
data_Age_Frequency <- fulldata %>%
  filter(!is.na(Age), Age > 10, Age < 65) %>%  # Filter valid ages
  group_by(Age) %>%  # Group by age
  summarise(frequency = n())  # Count the number of entries for each age

# Display the processed data
print(data_Age_Frequency)

# Plot: Reading frequency changes with age
ggplot(data = data_Age_Frequency, aes(x = Age, y = frequency)) +
  geom_line(color = "blue", linewidth = 1) +  # Use `linewidth` instead of `size`
  geom_point(color = "red", size = 2) +  # Keep `size` for points
  labs(
    x = "Age",
    y = "Reading Frequency",
    title = "How Reading Frequency Changes with Age"
  ) +
  theme_minimal() +
  scale_x_continuous(breaks = seq(10, 65, 5)) +  # Adjust x-axis ticks
  scale_y_continuous(expand = expansion(mult = c(0, 0.1)))  # Clean y-axis


```

Insights from the Visual:
Trend: Fiction engagement is highest between ages 25–35 and starts to decline after age 40.

ROW {}
-----------------------------------------------------------------------

### CHART 2

Let's see information on analysis about Genres:
```{r}
library(dplyr)
library(ggplot2)
library(tidyr)

# Process the data: Clean and split genres
fulldata <- fulldata %>%
  filter(!is.na(Age), Age > 10 & Age < 65) %>%
  mutate(genres = gsub("\\[|\\]|'", "", genres)) %>%
  separate_rows(genres, sep = ",\\s*")

# Group by genres and age groups
data_Genre_Frequency <- fulldata %>%
  mutate(Age_Group = cut(Age, breaks = seq(10, 65, 5), right = FALSE)) %>%
  group_by(genres, Age_Group) %>%
  summarise(frequency = n(), .groups = "drop")

# Focus on the top 10 most popular genres
top_genres <- data_Genre_Frequency %>%
  group_by(genres) %>%
  summarise(total_frequency = sum(frequency)) %>%
  arrange(desc(total_frequency)) %>%
  slice_head(n = 10) %>%
  pull(genres)

data_Genre_Frequency <- data_Genre_Frequency %>%
  filter(genres %in% top_genres)

# Plot: Reading frequency for top genres by age group
ggplot(data = data_Genre_Frequency, aes(x = Age_Group, y = frequency, fill = genres)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_brewer(palette = "Set2") +  # Use a visually appealing palette
  labs(
    x = "Age Group",
    y = "Reading Frequency",
    title = "Reading Frequency by Top Genres and Age Group"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),  # Rotate x-axis labels
    legend.position = "bottom"  # Move legend to the bottom for better layout
  )


```
-----------------------------------------------------------------------

What does this give to Book Producers?

- Target Popular Genres: Focus on genres with high frequency in key age groups (e.g., Fiction for 20-40).
- Expand Genre Appeal: Develop strategies for less popular genres in underrepresented age groups.
- Optimize Marketing:Tailor campaigns to age-group-specific genre preferences.



NEXT PLOT 

We separate the data having the only genre "Nonfiction" and afterwards only "fiction". 
We grouped the data set by Age and Calculate the mean rating for each age group and used this to plot a scatter plot.
Then we find the line of best fit. We calculate the slope with cor function, which calculates the correlation between age and mean rating for non-fiction genres.

We get a plot with linear regression line.

```{r}
# Filter rows containing 'Nonfiction' in the genres column
nonfiction_occurrences <- grep("Nonfiction", fulldata$genres)
nonfiction_rows <- fulldata[nonfiction_occurrences, ]

# Process Nonfiction data: Calculate mean age
nonfiction_age <- nonfiction_rows %>%
  filter(!is.na(Age), Age > 10, Age < 65) %>%  # Filter valid ages
  summarise(mean_age = mean(Age))  # Calculate mean age

# View the processed data
print(nonfiction_age)

# Plot: Distribution of Nonfiction Readers by Age
ggplot(data = nonfiction_rows, aes(x = Age)) +
  geom_histogram(binwidth = 5, fill = "darkgreen", alpha = 0.7, color = "black") +  # Histogram for age distribution
  labs(
    x = "Age",
    y = "Frequency",
    title = "Distribution of Ages for Nonfiction Readers"
  ) +
  theme_minimal()



```

The age of Nonfiction readers may have a minimal inverse relationship with the metric analyzed (e.g., frequency or rating).
In practical terms, as readers grow older, their engagement with the Nonfiction genre slightly declines, but the effect is not significant.

```{r}


# Filter rows containing 'Fiction' in the genres column
fiction_occurrences <- grep("Fiction", fulldata$genres)
fiction_rows <- fulldata[fiction_occurrences, ]

# Process Fiction data: Calculate frequency by age
fiction_age <- fiction_rows %>%
  filter(!is.na(Age), Age > 10, Age < 65) %>%  # Filter valid ages
  group_by(Age) %>%
  summarise(frequency = n(), .groups = "drop")  # Count occurrences of Fiction by age

# View the processed data
print(fiction_age)

# Plot: Frequency of Fiction Genre by Age
ggplot(data = fiction_age, aes(x = Age, y = frequency)) +
  geom_point(alpha = 0.5, color = "darkgreen") +  # Scatter plot
  geom_smooth(formula = y ~ x, method = "lm", se = TRUE, color = "darkgreen") +  # Linear trend line
  labs(
    x = "Age",
    y = "Frequency",
    title = "Frequency of Fiction Genre by Age"
  ) +
  ylim(0, max(fiction_age$frequency) * 1.1)  # Dynamically adjust y-axis limits

# Calculate correlation between Age and Frequency
corelation <- cor(fiction_age$Age, fiction_age$frequency)
corelation



```
Visual Insights:
Trend:
- The scatter plot shows a slight downward trend in Fiction frequency as age increases.
- Younger readers (20–30) seem to engage more with Fiction, with peak frequency near the mid-20s.
- Engagement drops gradually for older age groups (40–60).
Also:
- There is variability in Fiction readership across all age groups, but the trend suggests Fiction appeals slightly less to older readers.

Correlation Coefficient (-0.1503):
- The value indicates a very weak negative correlation.
- Fiction frequency slightly decreases as age increases, but the relationship is not strong enough to suggest age is a dominant factor.

- Age has a minimal impact on Fiction readership, meaning other factors (like genre preferences or access) might explain engagement more than age alone.
- Fiction appeals broadly across ages but shows higher engagement among younger (15-25) demographics.

Producers should:
- Target younger readers with campaigns and content aligned to their preferences.
Reinvigorate older audiences by introducing subgenres or formats (e.g., audiobooks) that cater to their interests.


```{r}
# Filter Fiction data for ages 15–40
fiction_occurrences <- grep("Fiction", fulldata$genres)
fiction_rows <- fulldata[fiction_occurrences, ]

fiction_15_40 <- fiction_rows %>%
  filter(!is.na(Age), Age >= 15 & Age <= 40) %>%
  group_by(Age) %>%
  summarise(frequency = n(), .groups = "drop")  # Frequency of Fiction by age

# Filter data for other genres for ages 15–40
other_genres <- fulldata %>%
  filter(!grepl("Fiction", genres), !is.na(Age), Age >= 15 & Age <= 40) %>%
  group_by(Age) %>%
  summarise(frequency = n(), .groups = "drop")  # Frequency of other genres by age

# Compare frequencies using a t-test
t_test <- t.test(fiction_15_40$frequency, other_genres$frequency, alternative = "greater")

# Print the t-test results
print(t_test)



```
p-value = 5.938e-09 < 0.05 : The p-value is less than 0.05, so we reject H₀.

The Welch Two Sample t-test was conducted to compare the engagement between Fiction and other genres among readers aged 15–40. The null hypothesis (H_0) assumed no significant difference in engagement, while the alternative hypothesis (H_1) proposed that Fiction has significantly higher engagement.
The test results showed a p-value of "5.938e-0", which is significantly less than 0.05.

Thus, we reject the null hypothesis and conclude that Fiction indeed exhibits significantly higher engagement compared to other genres in this demographic.

Conclusion:

- The lack of significant difference implies that Fiction is dominant among this demographic. 
- Producers should concentrate on Fiction sales, however we would recommend to also diversify offerings across genres and explore factors beyond age to identify potential areas for growth.

- We accept H1 Hypothesis, as we reject null hypothesis.

Business Side:
- Producers should concentrate on Fiction sales, but give a strike to other genres as well spread/popular between different ages.


